Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd

The construction of diagnostic plots adopts the core concept of grammar of graphics [@wickham2010layered] in ggplot2. In grammar of graphics, plots are not produced by calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but by the concept of stacked layers. Seeing plots as stacked layers gives analysts the freedom to compose plots with multiple elements on hand.

@@

This paper has set up a data object that can be used for diagnosing a complex optimisation procedure. The ideas were illustrated using the optimisers available for projection pursuit guided tour. Here the constraint is the orthonormality condition of the projection bases. The approach used broadly applies to other constrained optimisers.

<!-- something about the diagnostic plots  -->
Four diagnostic plots have been introduced to investigate the progression of the optimiser and its coverage in search space. The implementation of these visualisations is designed to be easy-to-use, each plot can be produced by supplying the data object to the corresponding function. More advanced users may decide to modify these basic graphs using the provided wrapper functions, or even build their own.

<!-- summarise what we have done and what might be done in the future. -->
Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \pkg{tourr} [@tourr] package; while the manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package. Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e. scagnostics, association, and information index [@laa2020using] and understand how the optimisers behave for index functions with different structures.

>>>>

<!-- The construction of diagnostic plots adopts the core concept of grammar of graphics [@wickham2010layered] in ggplot2. In grammar of graphics, plots are not produced by calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but by the concept of stacked layers. Seeing plots as stacked layers gives analysts the freedom to compose plots with multiple elements on hand. -->

@@

This paper has provided several visual diagnostics that can be used for diagnosing a complex optimisation procedure. The methods were illustrated using the optimisers available for projection pursuit guided tour. The methods are implemented in the package \pkg{ferrn}, and changes were made in the \pkg{tourr} package to generate the necessary data. This work complements with the work of @laa2020using and will be useful for researchers when designing new projection pursuit indexes.

This work might be considered an effort to bring transparency into algorithms. Although little attention is paid by algorithm developers to providing ways to output information during intermediate steps, this is an important component for enabling others to understand and diagnose the performance. Algorithms are an essential component of artificial intelligence that is used to make daily life easier. Interpretability of algorithms is important to guard against aspects like bias and inappropriate use. The data object underlying the visual diagnostics here is an example of what might be useful in algorithm development generally.

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex

This paper has set up a data object that can be used for diagnosing a
complex optimisation procedure. The ideas were illustrated using the
optimisers available for projection pursuit guided tour. Here the
constraint is the orthonormality condition of the projection bases. The
approach used broadly applies to other constrained optimisers.

Four diagnostic plots have been introduced to investigate the
progression of the optimiser and its coverage in search space. The
implementation of these visualisations is designed to be easy-to-use,
each plot can be produced by supplying the data object to the
corresponding function. More advanced users may decide to modify these
basic graphs using the provided wrapper functions, or even build their
own.

Most of the work in this project has been translated into code in two
packages: the collection of the data object is implemented in the
existing \CRANpkg{tourr} \citep{tourr} package; while the manipulation
and visualisation of the data object are implemented in the new
\pkg{ferrn} package. Equipped with handy tools to diagnose the
performance of optimisers, future work can extend the diagnostics to a
wider range of index functions, i.e.~scagnostics, association, and
information index \citep{laa2020using} and understand how the optimisers
behave for index functions with different structures.

>>>>

This paper has provided several visual diagnostics that can be used for
diagnosing a complex optimisation procedure. The methods were
illustrated using the optimisers available for projection pursuit guided
tour. The methods are implemented in the package \pkg{ferrn}, and
changes were made in the \pkg{tourr} package to generate the necessary
data. This work complements with the work of \citet{laa2020using} and
will be useful for researchers when designing new projection pursuit
indexes.

This work might be considered an effort to bring transparency into
algorithms. Although little attention is paid by algorithm developers to
providing ways to output information during intermediate steps, this is
an important component for enabling others to understand and diagnose
the performance. Algorithms are an essential component of artificial
intelligence that is used to make daily life easier. Interpretability of
algorithms is important to guard against aspects like bias and
inappropriate use. The data object underlying the visual diagnostics
here is an example of what might be useful in algorithm development
generally.
